---
title: "Coding Challenge 4"
author: "Brendan Willis"
date: "3/19/2023"
output: 
  html_document:
    code_folding: hide
editor_options:
  chunk_output_type: console
---
***1) Install the covidcast and DT packages using the following code. Note that you should run the following
code locally on your machine but you only need to install the packages once, not every time you knit
your assignment. Therefore, the following code should NOT be included in your knitted file.***



***2) Read in us_locations.csv and load the tidyverse package and covidcast package. Use the first
setup chunk at the top of your RMarkdown file to set the default chunk option message = FALSE,
which means messages will not be displayed in the HTML output. (Tip: you might want to try running
this with message = TRUE first and then changing it, to see the difference for yourself.) (1 pt)***

```{r, message=FALSE}
library(tidyverse)
library(DT)
library(covidcast)
states <- read_csv("/Users/brendanwillis/Downloads/us_locations.csv")
```


***3) The covidcast package allows you to query over 1 billion rows of surveillance data related to COVID-19
that are stored centrally in a database that is accessible through the Epidata API. Run the following
query to download confirmed COVID-19 case data as reported by the JHU Center for Systems Science
and Engineering (CSSE) dashboard. Note that the specified geo_values in the code below are FIPS
codes that correspond to Hampshire, Franklin and Middlesex counties in Massachusetts. (1 pt)***
```{r}
covid_cases <- covidcast_signal(data_source = "jhu-csse",
              signal = "confirmed_incidence_num",
              start_day = "2020-03-01", end_day = "2022-02-21",
              geo_values = c("25015", "25011", "25017"))
```


***4) Look at the resulting dataset in the local data viewer. What is the unit of analysis of this dataset? In
other words, what variables uniquely identify each set of random observations for a row? (2 pts)***

The unit of analysis of this dataset is a county on a day. The variables that uniquely identify this are geo_value and time_value. The variable geo_value indicates the county of each observation and time_value indicates the date that the observation is from.

***5) Use the group_by() and summarize() functions to compute the total number of confirmed COVID-19
cases reported in each of the three counties between March 1, 2020 and February 21, 2022. Print out
the results in your report. First, print out the results as standard R console output, as long as it is
clear which cumulative case count corresponds to each county. You should be able to complete this in
one chained dplyr-style set of commands. Second, display the resulting table as an interactive table
using DT::datatable() to show the results. (4 pts)***

```{r}

        
x <- covid_cases %>%
  group_by(geo_value)%>%
  summarize(sum(value))
  

head(x)
DT::datatable(x)
```

***6) Adapt the code you used in the previous question to print out the number of cases each county
had in all of 2021. Hint: you can use something like dplyr::filter(my_date_variable >
as.Date("2023-01-12") to filter rows that meet a specific criteria on a variable formatted as a Date
class. Show the results using DT::datatable(). (2 pts)***

```{r}
cases_2021 <- covid_cases %>% 
  filter(between(time_value, as.Date('2021-01-01'), as.Date('2021-12-31')))
table_2021 <- cases_2021 %>%
  group_by(geo_value)%>%
  summarize(sum(value))
DT::datatable(table_2021, rownames = FALSE, colnames = c("County", "Confirmed Cases"))
```

***7) To make the results in the previous question a bit more detailed and human-readable, perform the
following analyses (ideally, although not required to be, in one chained dplyr command) (4 pts):
a) use left_join() to join the us_locations dataset with the results from the previous quesion,
b) create a new variable called fraction_infected that computes the number of confirmed cases as
a fraction of the total population of the county,
c) print out the results in an interactive datatable with four columns (in this order): county name,
total confirmed infections, population, fraction infected.***

```{r}
count_name <- table_2021 %>%
  left_join(states, by=c("geo_value" = "fips"))%>%
  rename("case_count"="sum(value)")%>%
  mutate(fraction_infected = case_count/population)%>%
  select("location_name", "case_count", "population", "fraction_infected")

DT::datatable(count_name, rownames = FALSE, colnames = c("County", "Confirmed Infections","Population", "Fraction Infected"))
```



***8) Query the Epidata API for confirmed cases for all US counties in the month of January 2022 and save
this data as a new object in your workspace. (Tip: you can leave the geo_values argument unspecified
and it will return all counties.) (1 pt)***

```{r}
covid_cases_january <- covidcast_signal(data_source = "jhu-csse",
              signal = "confirmed_incidence_num",
              start_day = "2022-01-01", end_day = "2022-01-31"
              )
```

***9) Using a single chained set of dplyr commands, display an interactive datatable that shows all counties
in the US, but it should be pre-sorted so that the counties with the highest rate of confirmed
COVID-19 cases in January 2022 relative to their population size are at the top. (3 pts)***


```{r}
us_table2 <- covid_cases_january %>% 
  group_by(geo_value)%>%
  summarize(case_count=sum(value))%>%
  left_join(states, by=c("geo_value" = "fips"))%>%
  mutate(fraction_infected = case_count/population)%>%
  arrange(desc(fraction_infected))%>%
  select("location_name", "abbreviation","case_count", "population", "fraction_infected")%>%
  datatable(count_name, rownames = FALSE, colnames = c("County", "State", "Confirmed 
                                                       Infections","Population",
                                                       "Fraction Infected"))
us_table2
  
  
```

***10) Re-run the same analysis as in the question above but use group_by() and filter() so that only the
county with the highest fraction confirmed cases in its state in January 2022 is shown. The resulting
output should be an interactive table with 50 rows (only 10 shown at first by default), with only one
line for each state. The only row that should be present for each state is for the county which had the
highest rate in January 2022 for that particular state. (2 pts)***

```{r}
max_table <- covid_cases_january%>%
  group_by(geo_value)%>%
  summarize(case_count=sum(value))%>%
  left_join(states, by=c("geo_value" = "fips"))%>%
  mutate(fraction_infected = case_count/population)%>%
  group_by(abbreviation)%>%
  filter(fraction_infected == max(fraction_infected))%>%
  arrange(desc(fraction_infected))%>%
  select("location_name", "abbreviation","case_count", "population", "fraction_infected")%>%
  datatable(count_name, rownames = FALSE, colnames = c("County", "State", "Confirmed 
                                                       Infections","Population",
                                                       "Fraction Infected"))
max_table

  

```

